### 2025-04-09 backmatter/solutions.tex
T-1 day to initial submission, nearly complete draft up to citations and small comments.

@@

Let's examine the amplitude
\begin{align}
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (c) [above right=of mid] {\(k_n\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below=1cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid) -- (c),
            };
        \end{feynman}
    \end{tikzpicture}
    \,
    =
    \,\,\,
    \widetilde{\mc{M}}_{m\to n}(p_1,\cdots,p_m;\,k_1,\cdots,k_n)
    \,\,
    .
\end{align}
We will deal only with connected amplitudes in what follows.

Note that the only way to produce a \(\widetilde{\mc{M}}_{m \to n}\), at least perturbatively, is to use the cubic vertex of \phicubedsix{} to produce new outgoing lines from lower-order amplitudes:
\begin{align}
    \widetilde{\mc{M}}_{m\to n}(p_1,\cdots,p_m;\,k_1,\cdots,k_n)
    \,\,
    =
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (ap) [below right=.5cm and .5cm of a];
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (c) [above right=of mid] {\(k_n\)};
            \vertex (cp) [below=.7cm of c] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below=1.2cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                (ap) -- (c),
                (mid) -- (cp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    +
    \quad
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (cpp) [above right=.4cm and .4cm of mid];
            \vertex (c) [above right=.3cm and .6cm of cpp] {\(k_n\)};
            \vertex (cp) [right=.6cm of cpp] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below left=1.2cm and .3cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                % (cp) -- (cpp) -- (c)
                (cp) -- (cpp) -- (c),
                (mid) -- (cpp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    +
    \quad
    \cdots
    ,
    \label{eq:eikonal_diagram}
\end{align}
where the ellipses include the diagrams where \(k_n\) branches off of any of the external lines of the lower point amplitude.

\Eq{eikonal_diagram} indicates that the diagrams with \(n+m\) external lines can be produced by starting with diagrams with \(n+m-1\) external lines and including an additional propagator factor.
%
When attaching a new outgoing line to any of the ingoing particles, the propagator involves the difference of two momenta;
%
for example, the first diagram of \Eq{eikonal_diagram} is
\begin{align}
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (ap) [below right=.5cm and .5cm of a];
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (c) [above right=of mid] {\(k_n\)};
            \vertex (cp) [below=.7cm of c] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below=1.2cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                (ap) -- (c),
                (mid) -- (cp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    =
    \quad
    \frac{g}{(p_1 - k_n)^2 + m^2}
    \,\,\,\,
    \widetilde{\mc{M}}_{m \to n-1}(p_1 - k_n, p_2,\cdots, p_m;\,k_1,\cdots,k_{n-1})
    .
\end{align}
The additional factor due to the propagator is called an \vocab{eikonal factor}.

When instead attaching our new outgoing external line to the out-going particles, we instead have the sum of two momenta in the propagator
\begin{align}
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (cpp) [above right=.4cm and .4cm of mid];
            \vertex (c) [above right=.3cm and .6cm of cpp] {\(k_n\)};
            \vertex (cp) [right=.6cm of cpp] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below left=1.2cm and .3cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                % (cp) -- (cpp) -- (c)
                (cp) -- (cpp) -- (c),
                (mid) -- (cpp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    =
    \quad
    \frac{g}{(k_{n-1} + k_n)^2 + m^2}
    \,\,\,\,
    \widetilde{\mc{M}}_{m \to n-1}(p_1, p_2,\cdots, p_m;\,k_1,\cdots,k_{n-1}+k_n)
    .
\end{align}

\begin{exercise*}
    Show that
    \begin{align}
        \begin{tikzpicture}[baseline={(0,-0.1)}]
            \begin{feynman}
                \node[blob] (mid) {};
                \vertex (a) [above left=of mid] {\(p_1\)};
                \vertex (b) [below left=of mid] {\(p_m\)};
                \vertex (cpp) [above right=.4cm and .4cm of mid];
                \vertex (c) [above right=.3cm and .6cm of cpp] {\(k_n\)};
                \vertex (cp) [right=.6cm of cpp] {\(k_{n-1}\)};
                \vertex (d) [below right=of mid] {\(k_1\)};
                \node[below=1cm of a] (L) {\(\vdots\)};
                \node[below left=1.2cm and .3cm of c] (L) {\(\vdots\)};
                \diagram* {
                    (a) -- (mid) -- (b),
                    (d) -- (mid),
                    % (cp) -- (cpp) -- (c)
                    (cp) -- (cpp) -- (c),
                    (mid) -- (cpp),
                };
            \end{feynman}
        \end{tikzpicture}
        \!\!
        =
        \frac{g}{2 E_n E_{n-1}(1 - \cos\theta)}
        \widetilde{\mc{M}}_{m \to n-1}(p_1, p_2,\cdots, p_m;\,k_1,\cdots,k_{n-1}+k_n)
    \end{align}
    in the massless limit.
    %
    What is \(\theta\)?

@@

\begin{exercise*}
    What are the mass dimensions of \(\widetilde{\mc{M}}_{m\to n}\)?
    %
    Ensure that these mass dimensions make sense when combined with the formulae above.
\end{exercise*}

@@

Remember that the differential cross section for \(2 \to n\) scattering is given by
\begin{align}
    \dd \sigma_n
    =
    \frac{1}{u_\alpha 2 E_{p_1} 2 E_{p_2}} \le(2\pi\ri)^d
    \delta^{(d)}\le(k_1 + \cdots + k_n - p_1 - p_2\ri)
    \le|
        \widetilde{\mc{M}}_{2 \to n}
    \ri|
    \,
    \dd\widetilde{\Phi}_n
    ,
\end{align}
where \(u_\alpha\) is a ``relative velocity'' which depends only on the initial state, and
\begin{align}
    \dd\widetilde\Phi_n
    =
    \prod_{i=1}^n
    \frac{\dd^{d-1}k_i}{2 E_i \le(2\pi\ri)^{d-1}}
    .
\end{align}

@@

\begin{exercise*}
    Show that
    \begin{align}
        \abs{\Vec{k}'}
        =
        E'
        =
        E_n + E_{n-1} + \mc{O}\le(\theta^2\ri)
        ,
    \end{align}
    and use this to argue that, in the collinear limit, we can parameterize the energy dependence of the differential cross section into dependence on \(E'\) and on \(x\), where \(x\) is an \vocab{energy fraction} defined through
    \begin{align}
        E_n &= x E'
        \\
        E_{n-1} &= (1-x)E'
        .
    \end{align}
    Again, this will hold up to \(\mc{O}(\theta^2)\).
\end{exercise*}

@@

We now address \textbf{the collinear limit of the differential cross section}.
%
Let us write
\begin{align}
    \dd \sigma_n
    =
    \dd\sigma_{n-1}
    \,\,
    \times
    \frac{\le|\widetilde{\mc{M}}_{2\to n}\ri|^2}
    {\le|\widetilde{\mc{M}}_{2\to {n-1}}\ri|^2}
    \frac{\dd\widetilde{\Phi}_n}{\dd \widetilde\Phi_{n-1}}
    \
    ,
\end{align}
where we have used that \(\dd \widetilde\Phi_n\) contains a delta function that enforces \(k' = k_1 + k_2\).

Using all of the previous pieces we have derived, we can write
\begin{align}
    \frac{\le|\widetilde{\mc{M}}_{2\to n}\ri|^2}
    {\le|\widetilde{\mc{M}}_{2\to {n-1}}\ri|^2}
    &=
    \frac{1}{\le(E'\ri)^4}
    \frac{g^2}{4 x^2 (1-x)^2 \theta^4}
    \\
    \frac{\dd\widetilde{\Phi}_n}{\dd \widetilde\Phi_{n-1}}
    &\to
    \Omega_{d-3}
    \,\,
    \le(E'\ri)^{d-2}
    \,\,
    \frac{1}{(2\pi)^{d-1}}
    x^{d-3}
    (1-x)^{d-3}
    \,\,
    \frac{\dd \theta}{\theta^{3-d}}
    \,\,
    \dd x
    .
\end{align}

Therefore,
\begin{align}
    \dd\sigma_n
    =
    \dd \sigma_{n-1}
    \,\,
    \times
    \,\,
    \frac{g^2 E^{d-6}}{4\le(2\pi\ri)^{d-1}}
    \,\,
    \Omega_{d-3}
    \,\,
    \times
    \,\,
    \le(
        \dd x
        \,
        \le( x (1-x) \ri)^{d-5}
    \ri)
    \,\,
    \times
    \,\,
    \le(
        \frac{\dd\theta}{\theta^{7-d}}
    \ri)
    ,
\end{align}
and in \(\phicubedsix{}\),
\begin{align}
    \dd\sigma_n
    =
    \dd \sigma_{n-1}
    \,\,\,
    \dd x\,\, p(x)
    \,\,\,
    \frac{\dd \theta}{\theta}
    ,
\end{align}
where the \vocab{splitting function} \(p(x)\) is
\begin{align}
    p(x)
    =
    \frac{g^2\Omega_{d-3}}{4 \le(2 \pi\ri)^{d-1}}
    x^{d-5}(1-x)^{d-5}
    =
    2 a_g\, x(1-x)
\end{align}
in \(d=6\) dimensions, where \(a_g = g^2 \Omega_3 / 256 \pi^5 = g^2 / 128 \pi^3\) is the analogue of \(a_s = g_s^2 \Omega_2 / 64 \pi^3 = g_s^2 / \le(4 \pi\ri)^2\) in \gls{qcd}.

@@

In the standard model, and notably in \gls{qcd} where jets are major players, there is also a factor of \(\dd\theta/\theta\), so there is a collinear singularity.
%
However, some of the QCD splitting functions \(p_{kj\leftarrow i}(x)\) (where \(i, j\) and \(k\) denote ``partonic flavors'' such as ``quark'' or ``gluon'') also have singularities as \(x \to 0\);
%
these low-energy singularities are called \vocab{soft divergences}, and are important features of real-life collider physics experiments that do not appear in \phicubedsix.

@@

% ---------------------------------------
\Soln{Splitting Functions of the SM II: QCD}{massless-qcd}
% ---------------------------------------

% ---------------------------------------
\Soln{Mass vs. Angularity}{mass-vs-angularity}
% ---------------------------------------

@@

    % Draw arrows
    \draw [arrow] (start) -- (pro1);
    \draw [arrow] (pro1) -- (pro2);
    \draw [arrow] (pro2) -- (dec1);
    \draw [arrow] (dec1) -- node[]{yes}(pro3);
    \draw [arrow] (pro3) -- (dec2);
    \draw [arrow] (dec2) -- node[]{yes}(stop);
    % Draw special arrows
    \draw [arrow] (dec1) --  node[anchor=west, above=2pt ] {no} ++(-8,0) |- (pro1);    \draw [arrow] (dec2) --  node[anchor=east, above=2pt ] {no} ++(8,0) |- (pro1);
\end{tikzpicture}
}
% Caption
    \caption[The basic code flow of our simple parton shower algorithm.]{The basic code flow of the parton shower algorithm described in \App{partonshower}.
    %
    The veto algorithm is only used when considering MLL effects and is skipped at LL.
    %
    The computations of the algorithm are shown in orange, the decisions of the algorithm in green, and the start and stop of the algorithm in red.}
    \label{fig:ps_codeflow}
\end{figure}

@@

% ---------------------------------------
\Soln{Mellin Convolution Theorem}{mellinconvolution}
% ---------------------------------------
We want to show that
\begin{align}
    \mc M[f \mellinconvolution g](s) = \mc M[f](s) \mc M[g](1-s)
    =: \hat{f}(s) \hat{g}(1-s)
    ,
\end{align}
where
\(
    \hat{f}(s) := \mathcal{M}[f](s) = \int_0^\infty \dd x\, x^{s-1} f(x)
    ,
\)
and
\(
    (f \mellinconvolution g)(x)
    =
    \int_x^1 \frac{\dd y}{y}\, f(y) g(x/y)
    .
\)

@@

% ---------------------------------------
\Soln{Mellin Convolution Theorem II}{mellinconvolution2}
% ---------------------------------------
We want to show that
\begin{align}
    \mc M[f \circ_{\mc M} g](s) = \mc M[f](s) \mc M[g](1-s)
    =:
    \hat{f}(s) \hat{g}(1-s)
    \,,
\end{align}
where

\(
    (f(x) \circ_{\mc M} g(x))(x)
    =
    \int_0^\infty\,\dd\xi f(x\,\xi) g(\xi)
    .
\)

@@

%\sam{Notation for subjet EEC?}
%\begin{align}
%    \label{eq:subjet_eec_gr0}
%    \frac{1}{\sigma}\frac{\dd \Sigma_\text{EEC}}{\dd\chi}(\chi > 0, R)
%    =
%    \frac{3\,\ascf}{4}
%    \frac{\Theta\le(\chi > r^2/4\ri)}{\chi}
%    +
%    \mc{O}\le(\as^2\ri)
%\end{align}
%for \(\chi > 0\).
%%
%Including contributions at \(\chi = 0\), we have
%\begin{align}
%    \frac{1}{\sigma}\frac{\dd \Sigma_\text{EEC}}{\dd\chi}(\chi, R)
%    =
%    \delta(\chi)\le(1 - \frac{3\,\ascf}{2}\log(R/r)\ri)
%    +
%    \frac{3\,\ascf}{4}
%    \frac{\Theta\le(\chi > \frac{r^2}{4}\ri)}{\chi}
%    +
%    \mc{O}\le(\as^2\ri)
%    ,
%\end{align}
%where we have switched to normalizing with respect to the full cross section, allowing to determine the terms proportional to \(\delta(\chi)\) by enforcing the normalization constraint that \Eq{subjet_eec_gr} must integrate to one.
%%
%The terms proportional to \(\delta(\chi)\) encode, respectively, the no-emission probability -- for which there is only a single particle and only a single subjet -- and the possibility that a quark emits a gluon at an angle less than the subjet radius -- for which there are two particles but still a single subjet.

@@

In terms of these variables, we may write the NLO EEC in integral form (away from \(\chi = 0\), if we neglect contact terms with \(i=j\)) as
\begin{align}
    \label{eq:ee2hadrons_EEC}
    \frac{\dd\Sigma_\text{EEC}}{\dd\chi}
    (\chi>0)
    =
    2
    \ascf
    \sum_{i<j}
    \int_0^1 \dd x_i
    &
    \int_{1-x_i}^1 \dd x_j
    \frac{x_q^2 + x_{\overline{q}}^2}{(1-x_q)_+(1-x_{\overline{q}})_+}
    \,\frac{x_i}{2}
    \,\frac{x_j}{2}
    \\
    &
    \notag
    \delta\le(
        \chi
        -
        \le(
            \frac{1}{x_i} + \frac{1}{x_j} - \frac{1}{x_i x_j}
        \ri)
    \ri)
    ,
\end{align}
where we have added an overall factor of \(2\) on the right-hand-side to write the sum over non-contact contributions as \(\sum_{i \neq j} = 2 \sum_{i < j}\), and we have used that \(E_i/Q = x_i / 2\).

@@

% ---------------------------------------
\Soln{A Would-Be Particle-Level EWOC}{ee_ewoc}
% ---------------------------------------

>>>>

\begingroup
\renewcommand{\theequation}{\Aleph{chapter}.\arabic{equation}}

@@

Let's examine the amplitude
\begin{align}
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (c) [above right=of mid] {\(k_n\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below=1cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid) -- (c),
            };
        \end{feynman}
    \end{tikzpicture}
    \,
    =
    \,\,\,
    \widetilde{\mathcal{M}}_{m\to n}(p_1,\cdots,p_m;\,k_1,\cdots,k_n)
    \,\,
    .
\end{align}
We will deal only with connected amplitudes in what follows.

Note that the only way to produce a \(\widetilde{\mathcal{M}}_{m \to n}\), at least perturbatively, is to use the cubic vertex of \phicubedsix{} to produce new outgoing lines from lower-order amplitudes:
\begin{align}
    \widetilde{\mathcal{M}}_{m\to n}(p_1,\cdots,p_m;\,k_1,\cdots,k_n)
    \,\,
    =
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (ap) [below right=.5cm and .5cm of a];
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (c) [above right=of mid] {\(k_n\)};
            \vertex (cp) [below=.7cm of c] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below=1.2cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                (ap) -- (c),
                (mid) -- (cp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    +
    \quad
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (cpp) [above right=.4cm and .4cm of mid];
            \vertex (c) [above right=.3cm and .6cm of cpp] {\(k_n\)};
            \vertex (cp) [right=.6cm of cpp] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below left=1.2cm and .3cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                % (cp) -- (cpp) -- (c)
                (cp) -- (cpp) -- (c),
                (mid) -- (cpp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    +
    \quad
    \cdots
    ,
    \label{eq:eikonal_diagram}
\end{align}
where the ellipses include the diagrams where \(k_n\) branches off of any of the external lines of the lower point amplitude.

\Eq{eikonal_diagram} indicates that the diagrams with \(n+m\) external lines can be produced by starting with diagrams with \(n+m-1\) external lines and including an additional propagator factor.
%
When attaching a new outgoing line to any of the ingoing particles, the propagator involves the difference of two momenta;
%
for example, the first diagram of \Eq{eikonal_diagram} is
\begin{align}
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (ap) [below right=.5cm and .5cm of a];
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (c) [above right=of mid] {\(k_n\)};
            \vertex (cp) [below=.7cm of c] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below=1.2cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                (ap) -- (c),
                (mid) -- (cp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    =
    \quad
    \frac{g}{(p_1 - k_n)^2 + m^2}
    \,\,\,\,
    \widetilde{\mathcal{M}}_{m \to n-1}(p_1 - k_n, p_2,\cdots, p_m;\,k_1,\cdots,k_{n-1})
    .
\end{align}
The additional factor due to the propagator is called an \vocab{eikonal factor}.

\sam{TODO: fix metric to match rest of thesis}

When instead attaching our new outgoing external line to the out-going particles, we instead have the sum of two momenta in the propagator
\begin{align}
    \begin{tikzpicture}[baseline={(0,-0.1)}]
        \begin{feynman}
            \node[blob] (mid) {};
            \vertex (a) [above left=of mid] {\(p_1\)};
            \vertex (b) [below left=of mid] {\(p_m\)};
            \vertex (cpp) [above right=.4cm and .4cm of mid];
            \vertex (c) [above right=.3cm and .6cm of cpp] {\(k_n\)};
            \vertex (cp) [right=.6cm of cpp] {\(k_{n-1}\)};
            \vertex (d) [below right=of mid] {\(k_1\)};
            \node[below=1cm of a] (L) {\(\vdots\)};
            \node[below left=1.2cm and .3cm of c] (L) {\(\vdots\)};
            \diagram* {
                (a) -- (mid) -- (b),
                (d) -- (mid),
                % (cp) -- (cpp) -- (c)
                (cp) -- (cpp) -- (c),
                (mid) -- (cpp),
            };
        \end{feynman}
    \end{tikzpicture}
    \quad
    =
    \quad
    \frac{g}{(k_{n-1} + k_n)^2 + m^2}
    \,\,\,\,
    \widetilde{\mathcal{M}}_{m \to n-1}(p_1, p_2,\cdots, p_m;\,k_1,\cdots,k_{n-1}+k_n)
    .
\end{align}

\begin{exercise*}
    Show that
    \begin{align}
        \begin{tikzpicture}[baseline={(0,-0.1)}]
            \begin{feynman}
                \node[blob] (mid) {};
                \vertex (a) [above left=of mid] {\(p_1\)};
                \vertex (b) [below left=of mid] {\(p_m\)};
                \vertex (cpp) [above right=.4cm and .4cm of mid];
                \vertex (c) [above right=.3cm and .6cm of cpp] {\(k_n\)};
                \vertex (cp) [right=.6cm of cpp] {\(k_{n-1}\)};
                \vertex (d) [below right=of mid] {\(k_1\)};
                \node[below=1cm of a] (L) {\(\vdots\)};
                \node[below left=1.2cm and .3cm of c] (L) {\(\vdots\)};
                \diagram* {
                    (a) -- (mid) -- (b),
                    (d) -- (mid),
                    % (cp) -- (cpp) -- (c)
                    (cp) -- (cpp) -- (c),
                    (mid) -- (cpp),
                };
            \end{feynman}
        \end{tikzpicture}
        \!\!
        =
        \frac{g}{2 E_n E_{n-1}(1 - \cos\theta)}
        \widetilde{\mathcal{M}}_{m \to n-1}(p_1, p_2,\cdots, p_m;\,k_1,\cdots,k_{n-1}+k_n)
    \end{align}
    in the massless limit.
    %
    What is \(\theta\)?

@@

\begin{exercise*}
    What are the mass dimensions of \(\widetilde{\mathcal{M}}_{m\to n}\)?
    %
    Ensure that these mass dimensions make sense when combined with the formulae above.
\end{exercise*}

@@

Remember that the differential cross section for \(2 \to n\) scattering is given by
\begin{align}
    \dd \sigma_n
    =
    \frac{1}{u_\alpha 2 E_{p_1} 2 E_{p_2}} \le(2\pi\ri)^d
    \delta^{(d)}\le(k_1 + \cdots + k_n - p_1 - p_2\ri)
    \le|
        \widetilde{\mathcal{M}}_{2 \to n}
    \ri|
    \,
    \dd\widetilde{\Phi}_n
    ,
\end{align}
where \(u_\alpha\) is a ``relative velocity'' which depends only on the initial state, and
\begin{align}
    \dd\widetilde\Phi_n
    =
    \prod_{i=1}^n
    \frac{\dd^{d-1}k_i}{2 E_i \le(2\pi\ri)^{d-1}}
    .
\end{align}

@@

\begin{exercise*}
    Show that
    \begin{align}
        \abs{\Vec{k}'}
        =
        E'
        =
        E_n + E_{n-1} + \mathcal{O}\le(\theta^2\ri)
        ,
    \end{align}
    and use this to argue that, in the collinear limit, we can parameterize the energy dependence of the differential cross section into dependence on \(E'\) and on \(x\), where \(x\) is an \vocab{energy fraction} defined through
    \begin{align}
        E_n &= x E'
        \\
        E_{n-1} &= (1-x)E'
        .
    \end{align}
    Again, this will hold up to \(\mathcal{O}(\theta^2)\).
\end{exercise*}

@@

We now address \textbf{the collinear limit of the differential cross section}.
%
Let us write
\begin{align}
    \dd \sigma_n
    =
    \dd\sigma_{n-1}
    \,\,
    \times
    \frac{\le|\widetilde{\mathcal{M}}_{2\to n}\ri|^2}
    {\le|\widetilde{\mathcal{M}}_{2\to {n-1}}\ri|^2}
    \frac{\dd\widetilde{\Phi}_n}{\dd \widetilde\Phi_{n-1}}
    \
    ,
\end{align}
where we have used that \(\dd \widetilde\Phi_n\) contains a delta function that enforces \(k' = k_1 + k_2\).

Using all of the previous pieces we have derived, we can write
\begin{align}
    \frac{\le|\widetilde{\mathcal{M}}_{2\to n}\ri|^2}
    {\le|\widetilde{\mathcal{M}}_{2\to {n-1}}\ri|^2}
    &=
    \frac{1}{\le(E'\ri)^4}
    \frac{g^2}{4 x^2 (1-x)^2 \theta^4}
    \\
    \frac{\dd\widetilde{\Phi}_n}{\dd \widetilde\Phi_{n-1}}
    &\to
    \Omega_{d-3}
    \,\,
    \le(E'\ri)^{d-2}
    \,\,
    \frac{1}{(2\pi)^{d-1}}
    x^{d-3}
    (1-x)^{d-3}
    \,\,
    \frac{\dd \theta}{\theta^{3-d}}
    \,\,
    \dd x
    .
\end{align}

Therefore,
\begin{align}
    \dd\sigma_n
    =
    \dd \sigma_{n-1}
    \,\,
    \times
    \,\,
    \frac{g^2 E^{d-6}}{4\le(2\pi\ri)^{d-1}}
    \,\,
    \Omega_{d-3}
    \,\,
    \times
    \,\,
    \le(
        \dd x
        \,
        \le( x (1-x) \ri)^{d-5}
    \ri)
    \,\,
    \times
    \,\,
    \le(
        \frac{\dd\theta}{\theta^{7-d}}
    \ri)
    ,
\end{align}
and in \(\phicubedsix{}\),
\begin{align}
    \dd\sigma_n
    =
    \dd \sigma_{n-1}
    \,\,\,
    2 a_g
    \,\,
    \dd x\,\, p(x)
    \,\,\,
    \frac{\dd \theta}{\theta^{7-d}}
    =
    \dd \sigma_{n-1}
    \,\,\,
    2 a_g
    \,\,
    \dd x\,\, p(x)
    \,\,\,
    \frac{\dd \theta}{\theta}
    ,
\end{align}
where we have defined
\begin{align}
    a_g
    =
    \frac{g^2\Omega_{d-3}}{4 \le(2 \pi\ri)^{d-1}}
    =
    g^2 \Omega_3 / 256 \pi^5 = g^2 / 128 \pi^3
    \,,
\end{align}
where as before, the final equality holds in \(d=6\) dimensions -- this is the \phicubedsix{} analogue of \(a_s = g_s^2 \Omega_2 / 64 \pi^3 = g_s^2 / \le(4 \pi\ri)^2\) in \gls{qcd} -- and the \glslink{splittingfn}{\vocab{splitting function}} \(p(x)\) is
\begin{align}
    p(x)
    =
    \frac{g^2\Omega_{d-3}}{4 \le(2 \pi\ri)^{d-1}}
    x^{d-5}(1-x)^{d-5}
    =
    x(1-x)
    \,,
\end{align}
where the final equality holds again in \(d=6\) dimensions.

@@

In the standard model, and notably in \gls{qcd} where jets are major players, there is also a factor of \(\dd\theta/\theta\), so there is a collinear singularity.
%
However, some of the \gls{qcd} splitting functions \(P_{kj\leftarrow i}(x)\) (where \(i, j\) and \(k\) denote ``partonic flavors'' such as ``quark'' or ``gluon'') also have singularities as \(x \to 0\);
%
these low-energy singularities are called \vocab{soft divergences}, and are important features of real-life collider physics experiments that do not appear in \phicubedsix.

@@

@@

    % Draw arrows
    \draw [arrow] (start) -- (pro1);
    \draw [arrow] (pro1) -- (pro2);
    \draw [arrow] (pro2) -- (dec1);
    \draw [arrow] (dec1) -- node[]{yes}(pro3);
    \draw [arrow] (pro3) -- (dec2);
    \draw [arrow] (dec2) -- node[]{yes}(stop);
    % Draw special arrows
    \draw [arrow] (dec1) --  node[anchor=west, above=2pt ] {no} ++(-8,0) |- (pro1);    \draw [arrow] (dec2) --  node[anchor=east, above=2pt ] {no} ++(8,0) |- (pro1);
\end{tikzpicture}
}
% Caption
    \caption[The basic code flow of our simple parton shower algorithm.]{The basic code flow of the parton shower algorithm described in \Prob{partonshower}.
    %
    The veto algorithm is only used when considering MLL effects and is skipped at LL.
    %
    The computations of the algorithm are shown in orange, the decisions of the algorithm in green, and the start and stop of the algorithm in red.}
    \label{fig:ps_codeflow}
\end{figure}

@@

% ---------------------------------------
\Soln{Mellin Convolution Theorem}{mellinconvolution}
% ---------------------------------------
We want to show that
\begin{align}
    \mathcal M[f \mellinconvolution g](s) = \mathcal M[f](s) \mathcal M[g](1-s)
    =: \hat{f}(s) \hat{g}(1-s)
    ,
\end{align}
where
\(
    \hat{f}(s) := \mathcal{M}[f](s) = \int_0^\infty \dd x\, x^{s-1} f(x)
    ,
\)
and
\(
    (f \mellinconvolution g)(x)
    =
    \int_x^1 \frac{\dd y}{y}\, f(y) g(x/y)
    .
\)

@@

% ---------------------------------------
\Soln{Mellin Convolution Theorem II}{mellinconvolution2}
% ---------------------------------------
We want to show that
\begin{align}
    \mathcal M[f \circ_{\mathcal M} g](s) = \mathcal M[f](s) \mathcal M[g](1-s)
    =:
    \hat{f}(s) \hat{g}(1-s)
    \,,
\end{align}
where

\(
    (f(x) \circ_{\mathcal M} g(x))(x)
    =
    \int_0^\infty\,\dd\xi f(x\,\xi) g(\xi)
    .
\)

@@

%\sam{Notation for subjet EEC?}
%\begin{align}
%    \label{eq:subjet_eec_gr0}
%    \frac{1}{\sigma}\frac{\dd \Sigma_\text{EEC}}{\dd\chi}(\chi > 0, R)
%    =
%    \frac{3\,\ascf}{4}
%    \frac{\Theta\le(\chi > r^2/4\ri)}{\chi}
%    +
%    \mathcal{O}\le(\as^2\ri)
%\end{align}
%for \(\chi > 0\).
%%
%Including contributions at \(\chi = 0\), we have
%\begin{align}
%    \frac{1}{\sigma}\frac{\dd \Sigma_\text{EEC}}{\dd\chi}(\chi, R)
%    =
%    \delta(\chi)\le(1 - \frac{3\,\ascf}{2}\log(R/r)\ri)
%    +
%    \frac{3\,\ascf}{4}
%    \frac{\Theta\le(\chi > \frac{r^2}{4}\ri)}{\chi}
%    +
%    \mathcal{O}\le(\as^2\ri)
%    ,
%\end{align}
%where we have switched to normalizing with respect to the full cross section, allowing to determine the terms proportional to \(\delta(\chi)\) by enforcing the normalization constraint that \Eq{subjet_eec_gr} must integrate to one.
%%
%The terms proportional to \(\delta(\chi)\) encode, respectively, the no-emission probability -- for which there is only a single particle and only a single subjet -- and the possibility that a quark emits a gluon at an angle less than the subjet radius -- for which there are two particles but still a single subjet.

@@

In terms of these variables, we may write the NLO EEC in integral form (away from \(\chi = 0\), if we neglect contact terms with \(i=j\)) as
\begin{align}
    \label{eq:ee2hadrons_EEC}
    \frac{\dd\Sigma_\text{EEC}}{\dd\chi}
    (\chi>0)
    =
    4\,
    \ascf
    \sum_{i<j}
    \int_0^1 \dd x_i
    &
    \int_{1-x_i}^1 \dd x_j
    \frac{x_q^2 + x_{\overline{q}}^2}{(1-x_q)_+(1-x_{\overline{q}})_+}
    \,\frac{x_i}{2}
    \,\frac{x_j}{2}
    \\
    &
    \notag
    \delta\le(
        \chi
        -
        \le(
            \frac{1}{x_i} + \frac{1}{x_j} - \frac{1}{x_i x_j}
        \ri)
    \ri)
    ,
\end{align}
where we have added an overall factor of \(2\) on the right-hand-side to write the sum over non-contact contributions as \(\sum_{i \neq j} = 2 \sum_{i < j}\), and we have used that \(E_i/Q = x_i / 2\).

@@

\iffalse
% ---------------------------------------
\Soln{A Would-Be Particle-Level EWOC}{ee_ewoc}
% ---------------------------------------
